Bayesian Rate Estimation

### 2012-09-10 many.tex
Fixed typo in conclusion.  Added query on Jeffries prior.

To analyze our synthetic data set, we analytically marginalized over
the state flags (i.e.~cluster membership), using the likelihood in
Eq.~\eqref{eq:rate-shape-posterior}.  We did this to take advantage of
the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
requires all parameters to be in $\mathbb{R}$.  We applied a prior on
the shape parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$,
and an (approximately) Jeffreys prior on $r_0$,
\begin{equation}
  p\left( r_0 \right) = \frac{\sqrt{R_f}}{r_0}.
\end{equation}
\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps
    Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
    why the prior on $r_0$ should depend on $R_f$?]}  \will{Here is
  the calculation:
\begin{equation}
  p\left( r_0 \right) = \sqrt{\left\langle \left(\frac{\partial \log
      \mathcal{L}}{\partial r_0 } \right)^2 \right\rangle} = \sqrt{
    R_f \left\langle \left(\frac{\partial \log\hat{f}}{\partial r_0
    }\right)^2 \right\rangle } = \sqrt{R_f \left\langle \left(
    \frac{2}{r_0} - \frac{4 r_0}{r^2 + r_0^2} \right)^2 \right\rangle}
  = \sqrt{\frac{ 4 R_f}{3 r_0^2}} \propto \frac{\sqrt{R_f}}{r_0}.
\end{equation}
Note that the calculation of the prior uses the likelihood before
analytical marginalization over the flags, $f_i$.  We are ignoring the
fact that $R_f$ and $r_0$ are correlated.  In truth, the Jeffreys
prior would be the determinant of the matrix built out of terms like
$\partial \log\mathcal{L} / \partial \theta_i \times \partial
\log\mathcal{L} /\partial \theta_j$, but let's not get carried
away\ldots.  As another aside, note that we could have predicted the
$1/r_0$ dependence because $r_0$ has units, so the only unit-invariant
prior is to be flat in $\log r_0$; I'm don't have a similar argument
for the appearance of the $\sqrt{R_f}$ factor.}  (Note that this
factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
$1/\sqrt{R_f}$; we have verified that the priors on these parameters
are irrelevant to our results, as would be expected from the
measurement of $\sim 1000$ foreground stars.)

@@

Throughout this paper, we have assumed that the shape of the
foreground and background distributions is known, or at least can be
modeled with several additional parameters.  This is not necessarily
easy to do.  For example, in the case of gravitational-wave data
analysis, the shape of the foreground distribution of events may
depend on the details of a complex data-analysis pipeline as well as
the astrophysical source distribution, while the background event
distribution depends on data quality and may deviate significantly
from the simple Gaussian-noise behavior modeled in section
\ref{sec:GW-example}.  Several approaches have been developed to
accurately model both distributions, e.g., through the use of injected
signals to model the background \cite{ihope,CannonHannaKeppel:2012}.
However, this is a difficult problem (e.g., because of the need to
estimate the background at the very tails of the distribution), and
will require significant future work.

>>>>

To analyze our synthetic data set, we analytically marginalized over
the state flags (i.e.~cluster membership), using the likelihood in
Eq.~\eqref{eq:rate-shape-posterior}.  We did this to take advantage of
the \textit{emcee} sampler of \citet{ForemanMackey2012}, which
requires all parameters to be in $\mathbb{R}$.  We applied a prior on
the shape parameters that is flat in $\vec{x}_0$ and $\vec{\gamma}$,
and an (approximately) Jeffreys prior on $r_0$,
\begin{equation}
  p\left( r_0 \right) = \frac{\sqrt{R_f}}{r_0}.
\end{equation}
\ilya{[Why is this the Jeffreys prior (BTW, it's Jeffreys, or perhaps
    Jeffreys', but definitely not Jeffrey's ;) )?  It's not obvious
    why the prior on $r_0$ should depend on $R_f$?]}  \will{Here is
  the calculation:
\begin{equation}
  p\left( r_0 \right) = \sqrt{\left\langle \left(\frac{\partial \log
      \mathcal{L}}{\partial r_0 } \right)^2 \right\rangle} = \sqrt{
    R_f \left\langle \left(\frac{\partial \log\hat{f}}{\partial r_0
    }\right)^2 \right\rangle } = \sqrt{R_f \left\langle \left(
    \frac{2}{r_0} - \frac{4 r_0}{r^2 + r_0^2} \right)^2 \right\rangle}
  = \sqrt{\frac{ 4 R_f}{3 r_0^2}} \propto \frac{\sqrt{R_f}}{r_0}.
\end{equation}
}\ilya{Sorry, I still don't get it.  Isn't $\mathcal{L} = R_f \hat{f} + R_b \hat{b}$?  If so, how do you get the second equality above?}
\will{Note that the calculation of the prior uses the likelihood before
analytical marginalization over the flags, $f_i$.  We are ignoring the
fact that $R_f$ and $r_0$ are correlated.  In truth, the Jeffreys
prior would be the determinant of the matrix built out of terms like
$\partial \log\mathcal{L} / \partial \theta_i \times \partial
\log\mathcal{L} /\partial \theta_j$, but let's not get carried
away\ldots.  As another aside, note that we could have predicted the
$1/r_0$ dependence because $r_0$ has units, so the only unit-invariant
prior is to be flat in $\log r_0$; I'm don't have a similar argument
for the appearance of the $\sqrt{R_f}$ factor.}  (Note that this
factor of $\sqrt{R_f}$ cancels with the Jeffreys prior on the rate,
$1/\sqrt{R_f}$; we have verified that the priors on these parameters
are irrelevant to our results, as would be expected from the
measurement of $\sim 1000$ foreground stars.)

@@

Throughout this paper, we have assumed that the shape of the
foreground and background distributions is known, or at least can be
modeled with several additional parameters.  This is not necessarily
easy to do.  For example, in the case of gravitational-wave data
analysis, the shape of the foreground distribution of events may
depend on the details of a complex data-analysis pipeline as well as
the astrophysical source distribution, while the background event
distribution depends on data quality and may deviate significantly
from the simple Gaussian-noise behavior modeled in section
\ref{sec:GW-example}.  Several approaches have been developed to
accurately model both distributions, e.g., through the use of injected
signals to model the foreground distribution \cite{ihope,CannonHannaKeppel:2012}.
However, this is a difficult problem (e.g., because of the need to
estimate the background at the very tails of the distribution), and
will require significant future work.
